---
title: "Section 6 Mapping"
author: "Grace Wysocki"
date: "2/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##install.packages("maps")
library(maps)
library(tidyverse)
#> ── Attaching packages ─────────────────── tidyverse 1.3.1 ──
#> ✓ ggplot2 3.3.5     ✓ purrr   0.3.4
#> ✓ tibble  3.1.2     ✓ dplyr   1.0.7
#> ✓ tidyr   1.1.3     ✓ stringr 1.4.0
#> ✓ readr   2.0.2     ✓ forcats 0.5.1
#> ── Conflicts ────────────────────── tidyverse_conflicts() ──
#> x dplyr::filter() masks stats::filter()
#> x dplyr::lag()    masks stats::lag()
#> x purrr::map()    masks maps::map()
state_df <- ggplot2::map_data("state")
```

```{r}
ggplot(data = state_df,
            mapping = aes(x = long, y = lat,
                          group = group)) +
  geom_polygon() 
```

```{r}
##install.packages("usdata")
library(usdata)
state_stats
#> # A tibble: 51 x 24
#>    state abbr   fips pop2010 pop2000 homeownership multiunit
#>    <fct> <fct> <dbl>   <dbl>   <dbl>         <dbl>     <dbl>
#>  1 Alab… AL        1  4.78e6  4.45e6          71.1      15.5
#>  2 Alas… AK        2  7.10e5  6.27e5          64.7      24.6
#>  3 Ariz… AZ        4  6.39e6  5.13e6          67.4      20.7
#>  4 Arka… AR        5  2.92e6  2.67e6          67.7      15.2
#>  5 Cali… CA        6  3.73e7  3.39e7          57.4      30.7
#>  6 Colo… CO        8  5.03e6  4.30e6          67.6      25.6
#>  7 Conn… CT        9  3.57e6  3.41e6          69.2      34.6
#>  8 Dela… DE       10  8.98e5  7.84e5          73.6      17.7
#>  9 Dist… DC       11  6.02e5  5.72e5          43.5      61.7
#> 10 Flor… FL       12  1.88e7  1.60e7          69.7      30  
#> # … with 41 more rows, and 17 more variables: income <dbl>,
#> #   med_income <dbl>, poverty <dbl>, fed_spend <dbl>,
#> #   land_area <dbl>, smoke <dbl>, murder <dbl>,
#> #   robbery <dbl>, agg_assault <dbl>, larceny <dbl>,
#> #   motor_theft <dbl>, soc_sec <dbl>, nuclear <dbl>,
#> #   coal <dbl>, tr_deaths <dbl>, tr_deaths_no_alc <dbl>,
#> #   unempl <dbl>
```

```{r}
state_stats <- state_stats %>% mutate(state = str_to_lower(state))
state_full <- left_join(state_df, state_stats, by = c("region" = "state"))

ggplot(data = state_full, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = coal)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_void() +
  scale_fill_viridis_b()
```

Exercise 1 & 2:
```{r}
state_stats <- state_stats %>% mutate(state = str_to_lower(state))
state_full <- left_join(state_df, state_stats, by = c("region" = "state"))

ggplot(data = state_full, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = pop_change)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_void() +
  scale_fill_distiller(type = "div", direction = -1)
```
scale_fill_
fermenter is for binning continuous
brewer is for discrete colours
distiller is for continuous variable

6.2 Do We Need a Map?

Exercise 1. Think of a second example where, even though the data is spatial, the best graph to show a particular point would not be a map. (basically anything over time)

-covid cases per state over time
-births per state over time

Exercise 2. Refer back the United States examples that we completed in the earlier section. Choose a variable or two variables where a map makes the most sense to visualize that variable and explain your reasoning. Then, choose a variable or two variables where you might make a plot other than a map and explain your reasoning for why the map makes less sense for this variable.

6.3 Mapping other Areas

```{r}
active <- read_csv("https://raw.githubusercontent.com/iramler/stlawu_covid/main/slc_towns_active_cases.csv", n_max = 34)
#> Rows: 34 Columns: 330
#> ── Column specification ────────────────────────────────────
#> Delimiter: ","
#> chr   (1): NAME
#> dbl (329): Order, Population, Area, 1/18/2022, 1/14/2022...
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.
tcases <- read_csv("https://raw.githubusercontent.com/iramler/stlawu_covid/main/slc_towns_total_cases.csv", n_max = 34)
#> Rows: 34 Columns: 323
#> ── Column specification ────────────────────────────────────
#> Delimiter: ","
#> chr   (1): NAME
#> dbl (322): Order, Population, Area, 1/18/2022, 1/14/2022...
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.

active_long <- active %>% pivot_longer(5:ncol(active), names_to = "date",
                                       values_to = "cases")
## repeat for total cases
tcases_long <- tcases %>% pivot_longer(5:ncol(tcases), names_to = "date",
                                       values_to = "cases")

library(lubridate)
#> 
#> Attaching package: 'lubridate'
#> The following objects are masked from 'package:base':
#> 
#>     date, intersect, setdiff, union
covid_df <- left_join(tcases_long, active_long,
                      by = c("date", "Order", "NAME")) %>%
  mutate(date = mdy(date)) %>%
  rename(total_cases = cases.x,
         active_cases = cases.y) %>%
  mutate(total_cases = if_else(is.na(total_cases),
                               true = 0, false = total_cases),
         active_cases = if_else(is.na(active_cases),
                                      true = 0, false = active_cases))
  

covid_SLC <- covid_df %>% filter(NAME == "St. Lawrence County")
covid_sub <- covid_df %>% filter(NAME != "St. Lawrence County")

covid_SLC
covid_sub
```
Exercise 1:

```{r}
 ggplot(data = covid_SLC, aes(x = date, y = active_cases)) +
    geom_line()

```

```{r}
##install.packages("sf")
library(sf)
#> Linking to GEOS 3.8.1, GDAL 3.2.1, PROJ 7.2.1
shp <- read_sf("data/SLC_Civil_Boundaries_SHP/slc.shp") %>%
  st_transform(st_crs("+proj=longlat"))
shp
```

```{r}
ggplot(data = shp) +
  geom_sf() +
  theme_void()
```

```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date

ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() + scale_fill_distiller()
```

Exercise 2. Change the fill scale of the plot. Should you use an unordered, sequential, or diverging scale?
sequential
```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date

ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() + scale_fill_distiller()
```


Exercise 3. Change the colour scale so that active_cases are put into different bins with scale_fill_viridis_b(). What are some advantages and disadvantages of this?
```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date

ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() + scale_fill_viridis_b(n.breaks=10)
```


Exercise 4. Explore the ?geom_sf_text() function and add the actual number of cases to the subregions in the plot, as is done on the SLC website.

```{r}
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date

ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() + scale_fill_viridis_c() + geom_sf_text(aes(label = active_cases), color = "white")
```

6.4 Leaflet
```{r}
##install.packages("devtools")
library(devtools)
#> Loading required package: usethis
##devtools::install_github("li-wen-li/uszipcodes")
library(uszipcodes)
```

```{r}
beers <- read_csv("data/breweries.csv") 
#> Rows: 2407 Columns: 6
#> ── Column specification ────────────────────────────────────
#> Delimiter: ","
#> chr (5): brewery_name, type, address, website, state
#> dbl (1): state_breweries
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.
```

```{r}
raw_zip <- uszipcodes::get_zip(beers$address)
beers$Zip <- as.integer(uszipcodes::clean_zip(raw_zip))
#> Warning: NAs introduced by coercion

## only keep zip, lat, and long
zip_tab <- zip_table %>% dplyr::select(Zip, Latitude, Longitude)
beer_location <- inner_join(beers, zip_tab)
#> Joining, by = "Zip"
```

```{r}
content <- beer_location %>%
  mutate(popup = paste0('<a href =', beer_location$website, '>',
                        beer_location$brewery_name, '</a>'))
```

```{r}
library(leaflet)

beer_map <- leaflet(beer_location) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Wikimedia) %>% 
  addMarkers(lng = beer_location$Longitude, lat = beer_location$Latitude,
             clusterOptions = markerClusterOptions(),
             popup = content$popup)
```

Exercise 1:
removes zipcodes with no breweries and zipcodes that are messed up/incorrect

Exercise 2:
you have a lot of control of what you make the graph look like