---
title: "Untitled"
author: "Grace Wysocki"
date: "2/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Complete this quiz in a .Rmd file. To turn in the quiz, push both a .Rmd file and a knitted .html file to your GitHub site.

Statement of Integrity: Copy and paste the following statement and then sign your name (by typing it) on the line below.

“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.”

Grace Wysocki

Collaboration Reminder: You may not communicate with or work with anyone else on this quiz, but you may use any of our course materials or materials on the Internet.


Question 1 (20 points). Examine the alcohol.csv data set that you may have used as an introductory data set in STAT/DATA 234:
```{r}
library(tidyverse)
library(here)
alcohol_df <- read_csv(here("data/alcohol.csv"))
alcohol_df
```

The data set has the following variables on countries throughout the world:

country, the name of the country
beer_servings, the number of beer servings per person per year
spirit_servings, the number of spirit servings per person per year
wine_servings, the number of wine servings per person per year
total_litres_of_pure_alcohol, the number of litres of pure alcohol consumed per person per year

Construct a map of the beer_servings variable. You should choose an appropriate projection and change the default colour scheme of your map.

```{r}
library(tidyverse)
library(leaflet)
library(dplyr)
library(maps)
library(ggplot2)

map_data <- map_data("world")

map_edit <- rename(map_data, country = region)

map_full <- left_join(map_edit, alcohol_df, by = "country") 

#map_full <- map_data("world")
#ggplot(data = map_data, aes(x = long, y = lat, group = group) +
  #geom_map(data = map_data, map = map_data,
    #aes(x = long, y = lat, map_id = region),
    #color = "white", fill = "lightgray", size = 0.1) + 
     #coord_map(projection = "albers") +
  #theme_void() +
  #scale_fill_viridis_b()

  
```

```{r}
library(tidyverse)
library(leaflet)
library(dplyr)
library(maps)
library(ggplot2)

map_data <- map_data("world")

map_edit <- rename(map_data, country = region)

map_full <- left_join(map_edit, alcohol_df, by = "country") 

#ggplot() +
  #geom_map(
    #data = map_full, map = map_data,
    #aes(long, lat, map_id = country)
  #) + geom_polygon(aes(fill = beer_servings, color = ))

#ggplot(data = map_full) +
  #geom_sf(aes(fill = beer_servings)) + scale_fill_viridis(option="magma")
```
I tried.

Question 2 (5 points). Give an example of a question of interest about the alcohol data where you would want to construct a visualization other than a map to answer the question. What type of visualization would you construct instead?

What regions of the work consume the most alcohol?
This would be best to answer with a table accompanied by a graph such as a lollipop graph or a bar graph, where countries are grouped into regions to show ammounts.

