---
title: 'Section 7: Expressing Uncertainty and Variability'
author: "Grace Wysocki"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
#> ── Attaching packages ─────────────────── tidyverse 1.3.1 ──
#> ✓ ggplot2 3.3.5     ✓ purrr   0.3.4
#> ✓ tibble  3.1.2     ✓ dplyr   1.0.7
#> ✓ tidyr   1.1.3     ✓ stringr 1.4.0
#> ✓ readr   2.0.2     ✓ forcats 0.5.1
#> ── Conflicts ────────────────────── tidyverse_conflicts() ──
#> x dplyr::filter() masks stats::filter()
#> x dplyr::lag()    masks stats::lag()
pokemon_df <- read_csv("data/pokemon_full.csv")
#> New names:
#> * `` -> ...1
#> Rows: 721 Columns: 14
#> ── Column specification ────────────────────────────────────
#> Delimiter: ","
#> chr  (2): Name, Type
#> dbl (11): ...1, HP, Attack, Defense, Speed, SpAtk, SpDef...
#> lgl  (1): Legendary
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.
pokemon_height <- pokemon_df %>% 
  filter(Type %in% c("Bug", "Electric", "Fighting", "Flying",
                     "Grass", "Steel")) %>%
  group_by(Type) %>%
  summarise(avg_height = mean(height)) %>%
  mutate(Type = fct_reorder(Type, avg_height))
ggplot(data = pokemon_height, aes(x = Type, y = avg_height)) +
  geom_col() +
  coord_flip()
```

Exercise 1. What can’t we see from this graphic that would be useful?

We can"t see the variability/range of the heights in each type. We are also missing the counts of each type.

Exercise 2. Make a different plot that shows more relevant features about the underlying data.

```{r}
pokemon_small <- pokemon_df %>% 
  filter(Type %in% c("Bug", "Electric", "Fighting", "Flying",
                     "Grass", "Steel")) %>% 
  group_by(Type) %>%
  mutate(meanheight = mean(height, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(meanheight) %>% 
  mutate(Type = fct_inorder(Type))

ggplot(data = pokemon_small, aes(x = Type, y = height)) +
  geom_point(alpha = 0.3) +
  coord_flip() +
  geom_point(data = pokemon_height, aes(y = avg_height),  colour = "red")
  
```


Exercise 3. Though we are arguing that bar plots are not generally good to show summaries of continuous data, when might we want to make a bar plot of a summary of continuous data anyway?

- media has to consider the statistical knowledge of their audience
- sometimes you do not have the underlining data, only summary stats
- if the groupd have the same variability/distribution then it would be okay to just graph the averages

```{r}
##install.packages("openintro")
library(openintro)
#> Loading required package: airports
#> Loading required package: cherryblossom
#> Loading required package: usdata
data(mlb_players_18)
mlb_sum <- mlb_players_18 %>% group_by(position) %>%
  summarise(med_hr = median(HR)) %>%
  mutate(position = fct_reorder(position, med_hr))
ggplot(data = mlb_sum, aes(x = position, y = med_hr)) +
  geom_col() +
  coord_flip()
```

Exercise 4. “Fix” the previous plot to show the underlying variability in the number of homeruns for each player position by making a set of boxplots.

```{r}
mlb_players_18 <- mlb_players_18 %>% mutate(position = fct_reorder(position, HR, .fun = median))  %>% group_by(position) %>% 
  mutate(nplayers = n())

p1 <- ggplot(data = mlb_players_18, aes(x = position, y = HR)) +
  geom_boxplot() +
  coord_flip()
```


Exercise 5. Use the plotly package to give the sample size for each group when a user hovers over the boxes.

```{r}
library(plotly)
ggplotly(p1)

mlb_players_18 %>% group_by(position) %>% 
  summarise(nplayers = n())
```

7.3 Stat 113 survey

```{r}
statsurvey_df <- read_csv("data/stat113_survey.csv")
#> Rows: 5322 Columns: 21
#> ── Column specification ────────────────────────────────────
#> Delimiter: ","
#> chr (10): time_both, time_semester, time_year, Smoke, Ha...
#> dbl (11): Year, Hgt, Wgt, Sibs, Birth, Exercise, TV, GPA...
#> 
#> ℹ Use `spec()` to retrieve the full column specification for this data.
#> ℹ Specify the column types or set `show_col_types = FALSE` to quiet this message.
#> 
statsurvey_df %>%
  filter(!is.na(Tattoo)) %>%
  group_by(time_both, Tattoo) %>% 
  summarise(ncount = n()) %>% 
  ungroup() %>% 
  group_by(time_both) %>% 
  mutate(ntotal = sum(ncount)) %>% 
  ungroup() %>% 
  filter(Tattoo == "Yes") %>% mutate(prop = ncount/ntotal,
                                     se = sqrt(prop * (1-prop) / ntotal),
                                     l_se = prop - se,
                                     u_se = prop + se)

ggplot(data = statsurvey_df, aes(x = time_both,
                                 fill = Tattoo)) +
  geom_bar()
```




