---
title: "Section 2"
author: "grace wysocki"
date: "1/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.2.1
```{r}
## install.packages("palmerpenguins")
library(palmerpenguins)
library(tidyverse)
#> ── Attaching packages ─────────────────── tidyverse 1.3.1 ──
#> ✓ ggplot2 3.3.5     ✓ purrr   0.3.4
#> ✓ tibble  3.1.2     ✓ dplyr   1.0.7
#> ✓ tidyr   1.1.3     ✓ stringr 1.4.0
#> ✓ readr   2.0.2     ✓ forcats 0.5.1
#> ── Conflicts ────────────────────── tidyverse_conflicts() ──
#> x dplyr::filter() masks stats::filter()
#> x dplyr::lag()    masks stats::lag()
penguins
#> # A tibble: 344 x 8
#>    species island    bill_length_mm bill_depth_mm
#>    <fct>   <fct>              <dbl>         <dbl>
#>  1 Adelie  Torgersen           39.1          18.7
#>  2 Adelie  Torgersen           39.5          17.4
#>  3 Adelie  Torgersen           40.3          18  
#>  4 Adelie  Torgersen           NA            NA  
#>  5 Adelie  Torgersen           36.7          19.3
#>  6 Adelie  Torgersen           39.3          20.6
#>  7 Adelie  Torgersen           38.9          17.8
#>  8 Adelie  Torgersen           39.2          19.6
#>  9 Adelie  Torgersen           34.1          18.1
#> 10 Adelie  Torgersen           42            20.2
#> # … with 334 more rows, and 4 more variables:
#> #   flipper_length_mm <int>, body_mass_g <int>, sex <fct>,
#> #   year <int>

## single quantitative frequency plot
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_freqpoly()
#> `stat_bin()` using `bins = 30`. Pick better value with
#> `binwidth`.
#> Warning: Removed 2 rows containing non-finite values
#> (stat_bin).
```

```{r}
## single categorical barplot
ggplot(data = penguins, aes(x = species)) +
  geom_bar()
```

```{r}
## two quant scatterplot
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth() 
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
#> Warning: Removed 2 rows containing non-finite values
#> (stat_smooth).
#> Warning: Removed 2 rows containing missing values
#> (geom_point).
```
  
```{r}
## two cat heatmap
ggplot(data = penguins, aes(x = species, y = island)) +
  geom_bin_2d()
```
  
Review Exercises:

Exercise 1. In the above plots, we skipped the final “two variable” combination: one quantitative variable and one categorical variable. Create either a side-by-side boxplot or side-by-side violin plot of a quantitative and a categorical variable in the penguins data set.

Exercise 2. Modify the frequency plot made with geom_freqpoly() to use colour in two different ways:

change the colour of the line to any colour that R knows. A list of some colours can be found at this link.
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, color = species)) +
  geom_freqpoly(color = "blue") 
```

add a colour asethetic to make three different frequency lines, one for each species of penguin.
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, color = species)) +
  geom_freqpoly(aes(color = species))
```

Hint: recall that only aesthetic mappings (variables) go inside aes(). Of (a) or (b), which one is an aesthetic mapping?

Exercise 3. Thus far, we have only plotted one or two variables at a time. Recall that one way to construct a plot of three different variables is to use colour. Modify the scatterplot so that the points are coloured by species.
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  geom_smooth() 
```

Exercise 4. Recall that aes() aesthetics specified within the ggplot() function directly are called global aesthetics because every other GEOM will use them (unless specifically overridden) while aes() specified within a particular GEOM are called local aesthetics because only that particular GEOM will use them.

Modify the scatterplot and smoother so that

the points are coloured by species, but there is only one smoother instead of three.

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth() 
```

there are three different coloured smoothers (one for each species) but the points are all the same colour.

2.2.2 Facet

```{r}
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_histogram(colour = "black", fill = "white") +
  facet_wrap(~ species, ncol = 1)
#> `stat_bin()` using `bins = 30`. Pick better value with
#> `binwidth`.
#> Warning: Removed 2 rows containing non-finite values
#> (stat_bin).
#> ncol made the graphs stacked
```

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ species)
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
#> Warning: Removed 2 rows containing non-finite values
#> (stat_smooth).
#> Warning: Removed 2 rows containing missing values
#> (geom_point).
```

